### 2015-05-31 abstract.tex
Put things into git

\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,fancyhdr}
\usepackage[margin=1in,includehead]{geometry}

\begin{document}

  \begin{center}
    \Large\bf\boldmath Determining the Structure of Length-$k$ Steenrod Operations as $A(r)$-Modules
  \end{center}

  The Steenrod Algebra $\mathcal{A}$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.  Let $\mathcal{A}(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.  Consider the modules $L(k)$ spanned by sequences of Steenrod operations of length $k$.  Welcher proved that $L(k)$ is a free module over $\mathcal{A}(k-1)$.  We are interested in finding the structure of $L(k)$ as an $\mathcal{A}(r)$-module for any $r$.  We conjecture that $L(k)$ is built as an $\mathcal{A}(r)$-module out of $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, in the sense that it has an increasing filtration with quotients isomorphic to $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, and present partial results towards that claim.  In addition, we prove some interesting commutation relations in the Steenrod algebra relating to representations of Steenrod Algebra elements in Wood's $Z$-basis.

\end{document}

### 2015-05-31 poster/poster.tex
Put things into git

\documentclass[final]{beamer}

\mode<presentation>{
\usetheme{Bethlehem}
\usefonttheme[onlymath]{serif}
}
\usepackage{amsmath,amssymb,amsthm,fancyhdr,tikz,mathdots,bbm,tikz-cd,wrapfig,array,pigpen,enumerate,hyperref}
\usepackage[size=custom, width=121.9, height=91.44, scale=1.7]{beamerposter} %%%%Not final dims.

\newcommand{\fr}{\nobreak\mskip2mu\mathpunct{}\nonscript\mkern-\thinmuskip{:}\mskip3mu\relax}
\newcommand{\from}{\leftarrow}
\newcommand{\pathto}{\rightsquigarrow}
\newcommand{\inc}{\hookrightarrow}
\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Sq}{\mathrm{Sq}}
\newcommand{\mm}{/\!/\!}

\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}
\newtheorem{defn}{Definition}
\newtheorem{ex}{Example}

\title{Determining the Structure of Length-$k$ Steenrod Operations as $A(r)$-Modules}
\author{Benjamin Kraft}
\institute{Massachusetts Institute of Technology}

\begin{document}

  \begin{columns}[T]
    \begin{column}{0.245\textwidth} %%% LEFT SIDE
      \begin{block}{Cohomology Operations} %explain quickly, including stable operations
        Let $H^n(X;R)$ be the $n$th degree cohomology of the space $X$ with coefficients in the ring $R$.
        \begin{bulletitem}
          A \alert{cohomology operation} is a natural transformation of cohomology functors, that is, for $m$ and $n$ and rings $R$ and $S$, a map $\Theta\fr H^n(X;R)\to H^m(X;S)$ for each $X$, such that the diagram below commutes for each $f\fr X\to Y$.
        \end{bulletitem}
        \begin{center}
          \vspace{-0.4em}
          \begin{tikzcd}
            H^m(Y;R) \arrow{r}{\Theta} \arrow{d}{f^*} & H^n(Y;S) \arrow{d}{f^*} \\
            H^m(X;R) \arrow{r}{\Theta}                & H^n(X;S)
          \end{tikzcd}
          \vspace{-0.4em}
        \end{center}
        \begin{bulletitem}
          A \alert{stable cohomology operation} of degree $i$ is a sequence of cohomology operations $\Theta\fr H^n(X;R)\to H^{n+i}(X;S)$ for each $n$ which commute with the suspension isomorphism $\Sigma$:
        \end{bulletitem}
        \begin{center}
          \vspace{-0.4em}
          \begin{tikzcd}
            H^n(X;R) \arrow{r}{\Theta} \arrow{d}{\Sigma} & H^{n+i}(X;S) \arrow{d}{\Sigma} \\
            H^n(\Sigma X;R) \arrow{r}{\Theta}            & H^{n+i}(\Sigma X;S)
          \end{tikzcd}
          \vspace{-0.4em}
        \end{center}
      \end{block}
      \subsection{The Steenrod Algebra}

      The \alert{Steenrod Algebra} $\A$ is the algebra of stable cohomology operations from $\Z/2$ cohomology to itself.  It is generated by operations $\Sq^i$ of degree $i$ such that:
      \begin{bulletitemize}
        \item $\Sq^0$ is the identity
        \item $\Sq^n(x)=x^2$ if $x\in H^n(X;\Z/2)$
        \item $\Sq^n(x)=0$ if $x\in H^m(X;\Z/2)$, $m<n$
        \item $\Sq^n(xy)=\displaystyle\sum_{i+j=k} \Sq^i(x)\Sq^j(y)$.
      \end{bulletitemize}
      The entire algebra consists of the $\Sq^i$ and their compositions, which satisfy the \alert{Adem relations}:
      \[ \Sq^i\Sq^j = \sum_{k=0}^{\lfloor i/2 \rfloor} \binom{j-k-1}{i-2k}\Sq^{i+j-k}\Sq^k .\]
      Using the Adem relations, we can write any element in terms of the $\Sq^{2^i}$.  We let \alert{$\A(r)$} be the subalgebra generated by the $\Sq^{2^i}$ for $i\leq r$.
    \end{column}
    \begin{column}{0.495\textwidth} %%% CENTER
      \section{Bases of the Steenrod Algebra}
      \begin{columns}[T]
        \begin{column}{0.497\textwidth}
          \begin{block}{Adem (Admissible)}
            Using the Adem relations, any product $\Sq^i\Sq^j$ may be rewritten if $i<2j$.  Then the \alert{Adem basis} consists of $\Sq^{i_1}\cdots\Sq^{i_k}$, $i_j \geq 2i_{j+1}$.
            We may define \alert{$L(k)$} to be the quotient of the Steenrod Algebra by the ideal generated by all Adem basis elements of length greater than $k$.
          \end{block}
        \end{column}
        \begin{column}{0.497\textwidth}
          \begin{block}{Milnor (Dual)}
            Milnor found that the Steenrod Algebra is a Hopf algebra, with a commutative comultiplication.  Thus the dual Steenrod Algebra is commutative, and in fact is a polynomial algebra over elements $\xi_n$.  The duals of the monomials $\xi_1^{i_1}\xi_2^{i_2}\cdots$, denoted $\Sq(i_1,i_2,\ldots)$, form the \alert{Milnor basis}.
          \end{block}
        \end{column}
      \end{columns}
      \vspace{0.2em}
      \begin{columns}[T]
        \begin{column}{0.497\textwidth}
          \begin{block}{$P^s_t$}
            Let $P^s_t=\Sq(0,0,\ldots,0,2^s)$, where the $2^s$ is in the $t$th position.  Pick any ordering on the pairs $(s,t)$; products of $P^s_t$ in that order constitute a basis.  The $P^s_t$ derives some convenient properties from the dual basis, while also giving some flexibility in the choice of ordering, allowing it to describe ideals nicely.
          \end{block}
        \end{column}
        \begin{column}{0.497\textwidth}
          \begin{block}{Wood $Z$}
            The element $P^s_t$ has degree $2^s(2^t-1)$.  If we replace it by $\Sq^{2^s(2^t-1)}$, and use the reverse left lexicographic ordering on $(s + t,s)$, we get an analogue of the $P^s_t$ basis which is monomial in the $\Sq^i$; it describes certain ideals nicely like the $P^s_t$ bases, but is closer to the Adem basis.  It consists of subproducts of $\cdots\Sq^4\Sq^6\Sq^7\Sq^2\Sq^3\Sq^1$.
          \end{block}
        \end{column}
      \end{columns}
      \begin{section}{Structure of the Subalgebras}
        \begin{columns}[T]
          \begin{column}{0.33\textwidth}
            Welcher showed that $L(k)$ is a free $\A(k-1)$-module, so it is natural to consider the structure of $L(k)$ as an $\A(r)$-module for any $r$.  In particular, it appears that $L(k)$ is built up out of the ``quotient'' algebras $\A(r)\mm\A(r-k)$ (see below).
            \begin{block}{$\A(r)\mm\A(r-k)$}
              Consider $\A(r)$ as an $\A(r-k)$-module; note that we may also consider $\F_2$ to be an $\A(r-k)$-module, where the action of any element except $1$ is zero.  Then we may define \alert{$\A(r)\mm\A(r-k)$} to be the tensor algebra $\A(r)\otimes_{\A(r-k)}\F_2$.
              \begin{bulletitem}
                \alert{Intuitively}, $\A(r)\mm\A(r-k)$ is like $\A(r)$, but with any element which can be written as a product ending in $\Sq^{2^i}$, $i\leq r-k$, equal to zero.
              \end{bulletitem}
            \end{block}
          \end{column}
          \begin{column}{0.66\textwidth}
            \begin{block}{Small $\A(r)$}
              \begin{wrapfigure}{r}{4.2in}
                \vspace{-0.8em}
                \begin{center}
                  \hspace{-0.7in}\includegraphics[scale=2]{pics/A0.pdf}\\[1em]
                  \hspace{-0.7in}\includegraphics[scale=2]{pics/A1.pdf}
                  \vspace{0.2em}
                \end{center}
              \end{wrapfigure}
              $\A(0)$ is simple: it is spanned by $\Sq^0=1$ and $\Sq^1$, with $\Sq^1\Sq^1=0$.  It is shown as an $\A(0)$-module above right.  Each point represents an element of the (Adem) basis, and the thick solid line represents the (left) action of $\Sq^1$ (in this case, on $1$).  To represent an $\A(1)$ module, we use a thinner curved line for the (left) action of $\Sq^2$; $\A(1)$ is shown below right. So we know $\Sq^1\Sq^2=\Sq^3$ and $\Sq^2\Sq^3=\Sq^4\Sq^1+\Sq^5$.
            \end{block}
            \begin{wrapfigure}{R}{2in}
              \begin{center}
                \vspace{-1em}
                \hspace{-0.5in}\includegraphics[scale=2]{pics/L1-1.pdf}
              \end{center}
            \end{wrapfigure}
            \subsection{$L(1)$}
            \parbox{0.85\textwidth}{
              \begin{wrapfigure}{R}{2.5in}
                \begin{center}
                  \includegraphics[scale=2]{pics/A1-A0.pdf}
                \end{center}
              \end{wrapfigure}

              On the far right is shown the small-degree classes of $L(1)$ as an $\A(1)$-module; it continues periodically upwards.  Looking only at the thick lines, as an $\A(0)$-module, it is indeed free, with basis $\Sq^{2i}$.  As an $\A(1)$-module, it is not free, but it is still built up out of copies of $\A(1)\mm\A(0)$ (shown near right), in the sense that it has a filtration in which the filtration quotients are isomorphic to $\A(1)\mm\A(0)$, except for one containing only the single class $\Sq^0$.
            }
          \end{column}
        \end{columns}
      \end{section}
    \end{column}
    \begin{column}{0.245\textwidth} %%% RIGHT SIDE
      \vspace{1em}
      \subsection{$L(2)$}\\
      \begin{wrapfigure}{l}{6.3in}
        \begin{center}
          \vspace{-1em}
          \includegraphics[scale=2]{pics/L2-1.pdf}
          \vspace{-1em}
        \end{center}
      \end{wrapfigure}
      At right is shown $L(2)$ as a free module over $\A(1)$, as Welcher showed.  We can show that, indeed, $L(2)$ considered as an $\A(2)$ module has a filtration with filtration quotients in each dimension isomorphic to $\A(2)\mm\A(0)$, with the exception of a single class in every dimension, and a finite number of extra classes.

      \begin{block}{The general case}
        In the general case, we have shown, essentially, that the filtration exists, but not that it covers all of $L(k)$.  In particular, we prove the commutation identity that, if we let
        \[X_n = \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1},\]
        then
        \[X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\]
        This allows us to show that the top class $X_rX_{r-1}\cdots X_{r-k+1}$ of $\A(r)\mm\A(r-k)$, when multiplied by the basis elements $\Sq^{2^{r+1}m_1}\Sq^{2^rm_2}\cdots\Sq^{2^{r-k+2}m_k}$ for $m_1\geq m_2\geq\cdots m_{r-k}$, produces a linearly independent set.

        We hope to show that these filtrations in fact cover all but some set of classes which has, in the natural sense, lower dimensionality than $L(k)$; so far we have been able to show this only in specific cases, by finding the actual set of excluded classes in each case, but we hope to soon complete the general result.
      \end{block}
      \subsection{Acknowledgements}

      This project would not have been possible without the support of Mark Behrens, who proposed the problem and provided guidance along the way, and funding from the MIT Class of 1995 and Paul E. Gray Funds for UROP.

    \end{column}
  \end{columns}
\end{document}

### 2015-05-31 snippets/commutation-identity.tex
Put things into git

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}
  Recall that, in the Wood Z-basis, the top class of $A(n)\otimes_{A(n-1)}\F_2$ may be written as $\Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}$; denote this by $X_n$.

  \begin{lem}\label{lem:commutation-relation}
    For any positive integers $m$ and $n$,
    \[X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\]
  \end{lem}

  \begin{proof}
    First, we show the following identities by application of the Adem relations:
    \begin{prop}\label{prop:identity-1}
      For any positive integers $m$ and $n$,
      \[\Sq^{2^n-1}\Sq^{2^{n+1}m}=\Sq^{2^{n+1}m+2^n-1}.\]
    \end{prop}
    \begin{prop}\label{prop:identity-2}
      For any positive integers $k$, $m$, and $n$, with $k\leq n$,
      \[\Sq^{2^n-2^k}\Sq^{2^n-2^{k-1}+2^{n+1}m}=\Sq^{2^n-2^k+2^{n+1}m}\Sq^{2^n-2^{k-1}}+\Sq^{2^{n+1}-2^k}\Sq^{2^{n+1}m-2^{k-1}}.\]
    \end{prop}
    \begin{proof}[Proof of Proposition~\ref{prop:identity-1}]
      The Adem relations give us that
      \[\Sq^{2^n-1}\Sq^{2^{n+1}m} = \sum_{j=0}^{2^{n-1}-1}\binom{2^{n+1}m-1-j}{2^n-1-2j}\Sq^{2^{n+1}m+2^n-1-j}\Sq^j.\]
      Recall that in $\F_2$, $\binom{p}{q}$ is $1$ if and only if $p$ has a one in every position of its binary expansion that $q$ has a $1$.  Let $2^n>j>0$ and let $2^{\omega(j)}$ be the largest power of $2$ dividing $j$.  Then the binary expansion of $2^{n+1}m-1-j$ ends in a zero and then $\omega(j)$ ones, while that of $2^n-1-2j$ ends in a zero and then $\omega(j)+1$ ones, so the binomial coefficient in the Adem relation is zero.  Then the only term of the Adem relation that is nonzero is that when $j=0$, where $2^{n+1}m-1$ ends with $n+1$ ones, while $2^n-1$ consists of $n$ ones.
    \end{proof}
    \begin{proof}[Proof of Proposition~\ref{prop:identity-2}]
      We expand the first and last terms using the Adem relations, and show that they are identical except for the middle term. For convenience, let $d=2^{n+1}(m+1)-2^k-2^{k-1}$, the degree of the elements.
      \begin{align*}
        \Sq^{2^n-2^k}\Sq^{2^n-2^{k-1}+2^{n+1}m} &= \sum_{j=0}^{2^{n-1}-2^{k-1}} \binom{2^n-2^{k-1}+2^{n+1}m-1-j}{2^n-2^k-2j} \Sq^{d-j}\Sq^j \\
        \Sq^{2^{n+1}-2^k}\Sq^{2^{n+1}m-2^{k-1}} &=\sum_{j=0}^{2^n-2^{k-1}}\binom{2^{n+1}m-2^{k-1}-1-j}{2^{n+1}-2^k-2j}\Sq^{d-j}\Sq^j \\
      \end{align*}
      First, let $0\leq j\leq 2^{n-1}-2^{k-1}$; we will show the two binomial coefficients are equal.  We again consider the binary expansions.  The expansion of $2^n-2^k-2j$ is the same as that of $2^{n+1}-2^k-2j$, except that the leading one has been removed.  The corresponding bits of $2^n-2^{k-1}+2^{n+1}m-1-j$ and $2^{n+1}m-2^{k-1}-1-j$ are also zero and one, respectively, so the two binomial coefficients are the same.

      Now let $2^{n-1}-2^{k-1}<j<2^n-2^{k-1}$; we will show that the binomial coefficient
      \[\binom{2^{n+1}m-2^{k-1}-1-j}{2^{n+1}-2^k-2j}\]
      is zero.  Let $2^q$ be the next power of $2$ strictly larger than $2^n-2^{k-1}+j$, so $q\leq n-1$.  Then the $q$th position of the binary expansion will not satisfy the condition to make the binomial coefficient one, so it will be zero.

      Finally, if $j=2^n-2^{k-1}$, the binomial coefficient in the second formula has a zero on the bottom, thus is one, so we get the middle term of the claimed identity.
    \end{proof}

    Then in particular, by repeated application of the above identities, we may write
    \begin{align*}
      X_{n-1}\Sq^{2^{n+1}m} &= \Sq^{2^n-2^{n-1}}\Sq^{2^n-2^{n-2}}\cdots\Sq^{2^n-2}\Sq^{2^{n+1}m+2^n-1} \\
        &= \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}m+2^n-2}\Sq^{2^n-1} \\
        &\quad + \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}-2}\Sq^{2^{n+1}m-1}\\
        &\ \;\vdots \\
        &=\Sq^{2^{n+1}m}X_{n-1}\\
        &\quad + \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}-2}\Sq^{2^{n+1}m-1}\\
        &\quad + \cdots + \Sq^{2^n}\Sq^{2^{n+1}m-2^{n-1}}\Sq^{2^n-2^{n-2}}\cdots\Sq^{2^n-1}.
    \end{align*}
    Now all of the extra terms are products of (from left to right) an element of $\A(n-1)$, then an element of the form $\Sq^{2^{n+1}-2^k}$ for $k<n+1$, then some element of $\A$.  In particular, they are zero in $\A(n-1)\otimes_{\A(n)}\A$.  Then since $X_n$ is the top class of $\A(n)\otimes_{\A(n-1)}\F_2$, thus annihilates any element of $\A(n)$ that is not in $\A(n-1)$, $1\otimes x\mapsto X_nx$ is well-defined as a map from $\A(n-1)\otimes_{\A(n)}\A$ to $\A$, so the two sides are the same when multiplied by $X_n$:
    \[X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\]

  \end{proof}

\end{document}

### 2015-05-31 snippets/free-module.tex
Put things into git

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}
  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.

  \begin{prop}
    $\bar{L}(k)//A(n-k)$ is a free module over $A(n)//A(n-k)$, for any $n\geq k$.
  \end{prop}

  In particular, we will show that $\bar{L}(k)//A(n-k)$ has $A(n)//A(n-k)$-basis
  \[
    B=\{\Sq^I\mid I=(2^{n+1}j_1, 2^nj_2,\ldots, 2^{n-k+2}j_k), j_1\geq j_2\geq\cdots\geq j_k\geq 0\}.
  \]

  First, let us show that the module generated by $B$ is free.  Here we follow a similar argument of Welcher.  Let $t$ be the top class of $A(n)//A(n-k)$.  Then I claim that for any $A(n)//A(n-k)$-module $M$, the submodule generated by $m\in M$ is free if $tm\neq 0$.  % This depends on something in Margolis that I need to find.

  % Now I claim that the top class is
  % \[t=\Sq^{1+n 2^{n+1}}\Sq^{1+(n-1)2^n}\cdots\Sq^{1+(n-k+1)2^{n-k+2}} + t_0\Sq^1 + t_1\Sq^2 + \cdots + t_{n-k}\Sq^{2^{n-k}},\]
  % for some elements $t_i$.
  % By looking at Wood's Z-basis and the Adem relations it suffices to prove that
  % \[\Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}=\Sq^{1+n 2^{n+1}} + t_0\Sq^1 + t_1\Sq^2 + \cdots + t_{n-1}\Sq^{2^{n-1}},\]
  % again for some elements $t_i$.  To see this, note first that the two elements are in the right dimension.  Now since Wood's basis is a basis, the left side is nonzero in $A(n)//A(n-k)$, so it may be written as an element of $L(1)$ plus an element of $A(\Sq^i)_{i\leq n-1}$; the only possible nontrivial choice of the former is $\Sq^{1+n 2^{n+1}}$, and the latter by definition may be written in the form of the remaining terms.

  Then it suffices to show that $tB$ is linearly independent.  To do this, we consider only the top terms in the left lexicographic ordering, and show that no two are the same.  Following Welcher, we let $\equiv$ mean congruent modulo both lower terms and $A(\Sq^i)_{i\leq n-k}$.  In particular, we will show that if $K=(1+n 2^{n+1},1+(n-1)2^n,\ldots,1+(n-k+1)2^{n-k+2})$, and $\Sq^I\in B$, then $t\Sq^I\equiv\Sq^{I+K}$.  Clearly none of these are the same, so $tB$ will then be linearly independent.

  To show this, recall that in the Wood Z-basis, the top class $t$ is represented by the element
  \[\Sq^{2^{n+1}-2^n}\cdots\Sq^{2^{n+1}-1}\Sq^{2^n-2^{n-1}}\cdots\Sq^{2^{n-k+2}-2^{n-k+1}}\cdots\Sq^{2^{n-k+2}-1}.\]
  Then we get

  % \begin{lem} %not true!
  %   Let $t_{n,k}=\Sq^{1+n 2^{n+1}}\Sq^{1+(n-1)2^n}\cdots\Sq^{1+(n-k+1)2^{n-k+2}}$, with $t_{n,k}=1$ if $k\leq0$.  Then $t_{n,k}\Sq^{c 2^{n+1}} = \Sq^{c 2^{n+1} + 1 + n2^{n+1}}t_{n-1,k-1}$.
  % \end{lem}
  % \begin{proof}
  %   Fix $n-k$ and induct on $n$.  If $k=1$, we wish to show that
  %   \[\Sq^{1+n2^{n+1}}\Sq^{c2^{n+1}}\equiv\Sq^{1+n2^{n+1}+c2^{n+1}}

  % Next, let us show that $B$ has the appropriate size.  We use the Poincar\'e series $v(t)=\sum t^i \dim V_i$ of a graded vector space $V$; multiplication of tensor products corresponds to tensor products of graded vector spaces.  Let $l(t)$, $a(t)$, and $b(t)$ be the Poincar\'e series of $L(k)//A(n-k)$, $A(n)//A(n-k)$, and the space spanned by $B$, respectively.

  % By consideration of Wood's Z-basis, we get that
  % \[
  %   a(t) = \prod_{\substack{k+2\leq i\leq n+1\\ 0\leq j<i}} \left(1+t^{2^i-2^j}\right).
  % \]

\end{document}

### 2015-05-31 snippets/linearly-independent.tex
Put things into git

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}

  \begin{lem}\label{lem:linearly-independent}
    Taken as elements of $L(k)_{-\infty}$ as a module over $\A(n)\otimes_{\A(n-k)}\F_2$, the set
    \[\{X_nX_{n-1}\cdots X_{n-k+1}\Sq^I\mid I=(2^{n+1}m_1,2^nm_2,\ldots,2^{n-k+2}m_k), m_1\geq m_2\geq\cdots\geq m_{n-k}\}\]
    is linearly independent.
  \end{lem}
  \begin{proof}
    Note that by the commutation relation we have shown, the element $s=X_nX_{n-1}\cdots X_{n-k+1}\Sq^I$ may be rewritten as
    \[X_n\Sq^{2^{n+1}m_1}X_{n-1}\Sq^{2^nm_2}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_k}.\]
    Let $M_{k,n}$ be the subspace of $\A$ which is spanned by elements of the form $yz$ where $y\in L(k)$ and $z\in \A(n)$ with the degree of $z$ at least $1$.  If $b\in \A$, we will write $b+M_{k,n}$ for the subset of $\A$ consisting of elements of the form $b+m$ for $m\in M_{k,n}$.

    \begin{prop}\label{prop:m-thing}
      The element
      \[X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}}\]
      is in
      \[\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1} + M_{k,n-k}.\]
    \end{prop}
    \begin{proof}

    We proceed by induction on $k$ for fixed $n$.  If $k=1$, we know that $X_n\in M_{1,n-1}+\Sq^{n2^{n+1}+1}$, since the latter term is the only element of $L(1)$ in the right dimension.  Then since, as an $\A(n)$-module, $L(1)$ is periodic in degrees modulo $2^{n+1}$, $X_n\Sq^{2^{n+1}m}\in M_{1,n-1}+\Sq^{2^{n+1}(n+m)+1}$.

    Now suppose the proposition holds for some $k$; we will show it for $k+1$.  Let 
    \[t=X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}},\]
    let $t'=X_{n-k}\Sq^{2^{n-k+1}m_{n-k}}$, let
    \[b=\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1},\]
    and let $b'=\Sq^{2^{n-k+1}(m_{n-k}+n-k)+1}$.  Then by inductive hypothesis we know that $t\in b+M_{k,n-k}$ and by the base case we know that $t'\in b'+M_{1,n-k-1}$; we wish to show that $tt'\in bb'+M_{k+1,n-k-1}$.  To do this, it suffices to show that $bM_{1,n-k-1}$ and $M_{k,n-k}t'$ are contained in $M_{k+1,n-k-1}$; for simplicity we will show that this is true of the spanning sets we have constructed.

    First, let $y\in L(1)$, $z\in \A(n-k-1)$ with $z$ not in degree $0$, so that $yz\in M_{1,n-k-1}$.  Then by looking at degrees, $by$ is already in admissible form, so it is in $L(k+1)$, so $byz\in M_{k+1,n-k-1}$ as desired.  Since elements of the form $yz$ span $M_{1,n-k-1}$, we get that $bM_{1,n-k-1}\subset M_{k+1,n-k-1}$.

    Second, let $y\in L(k)$, $z\in \A(n-k)$ with $z$ not in degree $0$, so that $yz\in M_{k,n-k}$.  If $z\not\in\A(n-k-1)$, then $zX_{n-k}$ must be zero in $\A(n-k)$.  Then we may assume $z\in\A(n-k-1)$; we will show that both $yzb'$ and $yzM_{1,n-k-1}$ are contained in $M_{k+1,n-k-1}$.  Now if we write $zb'$ in admissible form, multiplying by $y$ will give an element already in admissible form by looking at degrees, so we need only show that $zb'\in M_{1,n-k-1}$.  If we multiply out using the Adem relations, this is clearly true, modulo showing that $zb'$ contains no terms of length $1$.  This is easy to show; by the Milnor basis, $z$ cannot end in a Steenrod square with degree a multiple of $2^{n-k}$, but $b'$ is a Steenrod square in a degree which is $1$ modulo $2^{n-k}$, so on multiplying the two, the length $1$ term must vanish, as desired.

    %maybe we can merge the following with the previous argument somehow?
    Finally, then, we must show that $yzM_{1,n-k-1}\subset M_{k+1,n-k-1}$; let $y'\in L(1)$ and $z'\in\A(n-k-1)$, so we must show that $yzy'z'\in M_{k+1,n-k-1}$.  Again looking at the Adem relations, we may write $zy'$ in the form $y''z''$ where $y''\in L(1)$ and $z''\in\A(n-k-1)$.  Then $z''z'\in \A(n-k-1)$ and is not in degree zero, and $yy''\in L(k+1)$, so $yy''z''z'\in M_{k+1,n-k-1}$.

    Then the induction is complete; all terms of $tt'-bb'$ are in $M_{k+1,n-k-1}$ as desired.
    \end{proof}

    Then since the elements in question may be written in the given form, their shortest terms are unique and distinct, so the terms are linearly independent.
  \end{proof}

\end{document}

### 2015-05-31 snippets/set-of-representatives.tex
Put things into git

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

\begin{document}

  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.

  \begin{prop}
    % There is a set of representatives for a basis of $\A(n)//\A(n-k)$ (in the sense of the above note), all of which lie in $\bar{L}(k)$.
    Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \bar{L}(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.  Equivalently, $\A(n)\subset \bar{L}(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
  \end{prop}

  \begin{proof}
    It suffices to prove the proposition on some basis for $\A(n)$.  Recall that in Wood's Z-basis of the Steenrod algebra, a basis for $\A(n)$ consists of all strings of squares selected from~\eqref{eq:woodz} and multiplied in the given order:
    \begin{equation} \label{eq:woodz}
      \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1} \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-1}\cdots\Sq^2\Sq^3\Sq^1
    \end{equation}

    We induct on the number of squares selected.  The empty product is $1$, which is in $\bar{L}(k)$, so we are done.  So now suppose the proposition holds on all elements of the Wood Z-basis of length at most $l-1$, and let $t$ be an element of length $l$.  Then, for some $m=2^j-2^i$, $i<j\leq n+1$, we may write $t=\Sq^m\left(x+yz\right)$ where $x$, $y$, and $z$ are as in the statement of the proposition.

    Now $\Sq^myz$ is clearly of the prescribed form, since $\Sq^my$ is in $\A$.  If $x\in\bar{L}(k-1)$, then $\Sq^m x\in \bar{L}(k)$ (since application of the Adem relations cannot increase the length of a monomial), so we may suppose $x\in L(k)$.

    Write $x$ in admissible form as $\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$.  If $\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$ is already in admissible form, then we have that $2^ki_k\leq 2^{k-1}i_{k-1}\leq\cdots\leq2i_1\leq m<2^{n+1}$, so in particular, $i_k<2^{n-k+1}$, so $i_k\in\A(\Sq^i)_{i\leq k}$, so in fact the entire product $\Sq^mx$ is in $\A(\Sq^i)_{i\leq k}$.  Then we may apply the Adem relations at least once, to get
    \[\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k} = a_0\Sq^{m+i_1}\Sq^{i_2}\cdots\Sq^{i_k} + \sum_{1\leq j\leq \frac{m}2}a_j\Sq^{m+i_1-j}\Sq^j\Sq^{i_2}\cdots\Sq^{i_k};\]
    the first term we know is in $L(k)$.  For each of the remaining terms, if we may repeat the same process, ignoring the first factor; either the product is in $\A(\Sq^i)_{i\leq k}$, or it may be further reduced by the Adem relations.  Finally, if we apply the Adem relations $k$ times, the rightmost factor must be $\Sq^i$ for $i\leq \frac{m}{2^k}<2^{n-k+1}$, so similarly the product lies in the ideal $\A(\Sq^i)_{i\leq k}$.  Then $\Sq^mx$ may be written in the desired form, so we are done.
  \end{proof}

\end{document}

### 2015-05-31 paper.tex
Dump snippets into paper

\documentclass{article}

\usepackage{amsmath,amssymb,amsfonts,amsthm}

\newcommand{\A}{\mathcal{A}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Sq}{\mathrm{Sq}}

\newtheorem{prop}{Proposition}
\newtheorem{lem}{Lemma}

%TODO: frontmatter
\title{Determining the Structure of Length-$k$ Steenrod Operations as $A(r)$-Modules}
\author{Benjamin Kraft}

\begin{document}

\begin{abstract}
  The Steenrod Algebra $\mathcal{A}$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.  Let $\mathcal{A}(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.  Consider the modules $L(k)$ spanned by sequences of Steenrod operations of length $k$.  Welcher proved that $L(k)$ is a free module over $\mathcal{A}(k-1)$.  We are interested in finding the structure of $L(k)$ as an $\mathcal{A}(r)$-module for any $r$.  We conjecture that $L(k)$ is built as an $\mathcal{A}(r)$-module out of $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, in the sense that it has an increasing filtration with quotients isomorphic to $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, and present partial results towards that claim.  In addition, we prove some interesting commutation relations in the Steenrod algebra relating to representations of Steenrod Algebra elements in Wood's $Z$-basis.
\end{abstract}

\section{The Commutation Relation}

  Recall that, in the Wood Z-basis, the top class of $A(n)\otimes_{A(n-1)}\F_2$ may be written as $\Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}$; denote this by $X_n$.

  \begin{lem}\label{lem:commutation-relation}
    For any positive integers $m$ and $n$,
    \[X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\]
  \end{lem}

  \begin{proof}
    First, we show the following identities by application of the Adem relations:
    \begin{prop}\label{prop:identity-1}
      For any positive integers $m$ and $n$,
      \[\Sq^{2^n-1}\Sq^{2^{n+1}m}=\Sq^{2^{n+1}m+2^n-1}.\]
    \end{prop}
    \begin{prop}\label{prop:identity-2}
      For any positive integers $k$, $m$, and $n$, with $k\leq n$,
      \[\Sq^{2^n-2^k}\Sq^{2^n-2^{k-1}+2^{n+1}m}=\Sq^{2^n-2^k+2^{n+1}m}\Sq^{2^n-2^{k-1}}+\Sq^{2^{n+1}-2^k}\Sq^{2^{n+1}m-2^{k-1}}.\]
    \end{prop}
    \begin{proof}[Proof of Proposition~\ref{prop:identity-1}]
      The Adem relations give us that
      \[\Sq^{2^n-1}\Sq^{2^{n+1}m} = \sum_{j=0}^{2^{n-1}-1}\binom{2^{n+1}m-1-j}{2^n-1-2j}\Sq^{2^{n+1}m+2^n-1-j}\Sq^j.\]
      Recall that in $\F_2$, $\binom{p}{q}$ is $1$ if and only if $p$ has a one in every position of its binary expansion that $q$ has a $1$.  Let $2^n>j>0$ and let $2^{\omega(j)}$ be the largest power of $2$ dividing $j$.  Then the binary expansion of $2^{n+1}m-1-j$ ends in a zero and then $\omega(j)$ ones, while that of $2^n-1-2j$ ends in a zero and then $\omega(j)+1$ ones, so the binomial coefficient in the Adem relation is zero.  Then the only term of the Adem relation that is nonzero is that when $j=0$, where $2^{n+1}m-1$ ends with $n+1$ ones, while $2^n-1$ consists of $n$ ones.
    \end{proof}
    \begin{proof}[Proof of Proposition~\ref{prop:identity-2}]
      We expand the first and last terms using the Adem relations, and show that they are identical except for the middle term. For convenience, let $d=2^{n+1}(m+1)-2^k-2^{k-1}$, the degree of the elements.
      \begin{align*}
        \Sq^{2^n-2^k}\Sq^{2^n-2^{k-1}+2^{n+1}m} &= \sum_{j=0}^{2^{n-1}-2^{k-1}} \binom{2^n-2^{k-1}+2^{n+1}m-1-j}{2^n-2^k-2j} \Sq^{d-j}\Sq^j \\
        \Sq^{2^{n+1}-2^k}\Sq^{2^{n+1}m-2^{k-1}} &=\sum_{j=0}^{2^n-2^{k-1}}\binom{2^{n+1}m-2^{k-1}-1-j}{2^{n+1}-2^k-2j}\Sq^{d-j}\Sq^j \\
      \end{align*}
      First, let $0\leq j\leq 2^{n-1}-2^{k-1}$; we will show the two binomial coefficients are equal.  We again consider the binary expansions.  The expansion of $2^n-2^k-2j$ is the same as that of $2^{n+1}-2^k-2j$, except that the leading one has been removed.  The corresponding bits of $2^n-2^{k-1}+2^{n+1}m-1-j$ and $2^{n+1}m-2^{k-1}-1-j$ are also zero and one, respectively, so the two binomial coefficients are the same.

      Now let $2^{n-1}-2^{k-1}<j<2^n-2^{k-1}$; we will show that the binomial coefficient
      \[\binom{2^{n+1}m-2^{k-1}-1-j}{2^{n+1}-2^k-2j}\]
      is zero.  Let $2^q$ be the next power of $2$ strictly larger than $2^n-2^{k-1}+j$, so $q\leq n-1$.  Then the $q$th position of the binary expansion will not satisfy the condition to make the binomial coefficient one, so it will be zero.

      Finally, if $j=2^n-2^{k-1}$, the binomial coefficient in the second formula has a zero on the bottom, thus is one, so we get the middle term of the claimed identity.
    \end{proof}

    Then in particular, by repeated application of the above identities, we may write
    \begin{align*}
      X_{n-1}\Sq^{2^{n+1}m} &= \Sq^{2^n-2^{n-1}}\Sq^{2^n-2^{n-2}}\cdots\Sq^{2^n-2}\Sq^{2^{n+1}m+2^n-1} \\
        &= \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}m+2^n-2}\Sq^{2^n-1} \\
        &\quad + \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}-2}\Sq^{2^{n+1}m-1}\\
        &\ \;\vdots \\
        &=\Sq^{2^{n+1}m}X_{n-1}\\
        &\quad + \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-4}\Sq^{2^{n+1}-2}\Sq^{2^{n+1}m-1}\\
        &\quad + \cdots + \Sq^{2^n}\Sq^{2^{n+1}m-2^{n-1}}\Sq^{2^n-2^{n-2}}\cdots\Sq^{2^n-1}.
    \end{align*}
    Now all of the extra terms are products of (from left to right) an element of $\A(n-1)$, then an element of the form $\Sq^{2^{n+1}-2^k}$ for $k<n+1$, then some element of $\A$.  In particular, they are zero in $\A(n-1)\otimes_{\A(n)}\A$.  Then since $X_n$ is the top class of $\A(n)\otimes_{\A(n-1)}\F_2$, thus annihilates any element of $\A(n)$ that is not in $\A(n-1)$, $1\otimes x\mapsto X_nx$ is well-defined as a map from $\A(n-1)\otimes_{\A(n)}\A$ to $\A$, so the two sides are the same when multiplied by $X_n$:
    \[X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\]

  \end{proof}

\section{Representatives for \boldmath$\A(n)//\A(n-k)$}

  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.

  \begin{prop}
    % There is a set of representatives for a basis of $\A(n)//\A(n-k)$ (in the sense of the above note), all of which lie in $\bar{L}(k)$.
    Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \bar{L}(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.  Equivalently, $\A(n)\subset \bar{L}(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
  \end{prop}

  \begin{proof}
    It suffices to prove the proposition on some basis for $\A(n)$.  Recall that in Wood's Z-basis of the Steenrod algebra, a basis for $\A(n)$ consists of all strings of squares selected from~\eqref{eq:woodz} and multiplied in the given order:
    \begin{equation} \label{eq:woodz}
      \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1} \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-1}\cdots\Sq^2\Sq^3\Sq^1
    \end{equation}

    We induct on the number of squares selected.  The empty product is $1$, which is in $\bar{L}(k)$, so we are done.  So now suppose the proposition holds on all elements of the Wood Z-basis of length at most $l-1$, and let $t$ be an element of length $l$.  Then, for some $m=2^j-2^i$, $i<j\leq n+1$, we may write $t=\Sq^m\left(x+yz\right)$ where $x$, $y$, and $z$ are as in the statement of the proposition.

    Now $\Sq^myz$ is clearly of the prescribed form, since $\Sq^my$ is in $\A$.  If $x\in\bar{L}(k-1)$, then $\Sq^m x\in \bar{L}(k)$ (since application of the Adem relations cannot increase the length of a monomial), so we may suppose $x\in L(k)$.

    Write $x$ in admissible form as $\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$.  If $\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$ is already in admissible form, then we have that $2^ki_k\leq 2^{k-1}i_{k-1}\leq\cdots\leq2i_1\leq m<2^{n+1}$, so in particular, $i_k<2^{n-k+1}$, so $i_k\in\A(\Sq^i)_{i\leq k}$, so in fact the entire product $\Sq^mx$ is in $\A(\Sq^i)_{i\leq k}$.  Then we may apply the Adem relations at least once, to get
    \[\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k} = a_0\Sq^{m+i_1}\Sq^{i_2}\cdots\Sq^{i_k} + \sum_{1\leq j\leq \frac{m}2}a_j\Sq^{m+i_1-j}\Sq^j\Sq^{i_2}\cdots\Sq^{i_k};\]
    the first term we know is in $L(k)$.  For each of the remaining terms, if we may repeat the same process, ignoring the first factor; either the product is in $\A(\Sq^i)_{i\leq k}$, or it may be further reduced by the Adem relations.  Finally, if we apply the Adem relations $k$ times, the rightmost factor must be $\Sq^i$ for $i\leq \frac{m}{2^k}<2^{n-k+1}$, so similarly the product lies in the ideal $\A(\Sq^i)_{i\leq k}$.  Then $\Sq^mx$ may be written in the desired form, so we are done.
  \end{proof}

  \section{Linear Independence}

  \begin{lem}\label{lem:linearly-independent}
    Taken as elements of $L(k)_{-\infty}$ as a module over $\A(n)\otimes_{\A(n-k)}\F_2$, the set
    \[\{X_nX_{n-1}\cdots X_{n-k+1}\Sq^I\mid I=(2^{n+1}m_1,2^nm_2,\ldots,2^{n-k+2}m_k), m_1\geq m_2\geq\cdots\geq m_{n-k}\}\]
    is linearly independent.
  \end{lem}
  \begin{proof}
    Note that by the commutation relation we have shown, the element $s=X_nX_{n-1}\cdots X_{n-k+1}\Sq^I$ may be rewritten as
    \[X_n\Sq^{2^{n+1}m_1}X_{n-1}\Sq^{2^nm_2}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_k}.\]
    Let $M_{k,n}$ be the subspace of $\A$ which is spanned by elements of the form $yz$ where $y\in L(k)$ and $z\in \A(n)$ with the degree of $z$ at least $1$.  If $b\in \A$, we will write $b+M_{k,n}$ for the subset of $\A$ consisting of elements of the form $b+m$ for $m\in M_{k,n}$.

    \begin{prop}\label{prop:m-thing}
      The element
      \[X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}}\]
      is in
      \[\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1} + M_{k,n-k}.\]
    \end{prop}
    \begin{proof}

    We proceed by induction on $k$ for fixed $n$.  If $k=1$, we know that $X_n\in M_{1,n-1}+\Sq^{n2^{n+1}+1}$, since the latter term is the only element of $L(1)$ in the right dimension.  Then since, as an $\A(n)$-module, $L(1)$ is periodic in degrees modulo $2^{n+1}$, $X_n\Sq^{2^{n+1}m}\in M_{1,n-1}+\Sq^{2^{n+1}(n+m)+1}$.

    Now suppose the proposition holds for some $k$; we will show it for $k+1$.  Let 
    \[t=X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}},\]
    let $t'=X_{n-k}\Sq^{2^{n-k+1}m_{n-k}}$, let
    \[b=\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1},\]
    and let $b'=\Sq^{2^{n-k+1}(m_{n-k}+n-k)+1}$.  Then by inductive hypothesis we know that $t\in b+M_{k,n-k}$ and by the base case we know that $t'\in b'+M_{1,n-k-1}$; we wish to show that $tt'\in bb'+M_{k+1,n-k-1}$.  To do this, it suffices to show that $bM_{1,n-k-1}$ and $M_{k,n-k}t'$ are contained in $M_{k+1,n-k-1}$; for simplicity we will show that this is true of the spanning sets we have constructed.

    First, let $y\in L(1)$, $z\in \A(n-k-1)$ with $z$ not in degree $0$, so that $yz\in M_{1,n-k-1}$.  Then by looking at degrees, $by$ is already in admissible form, so it is in $L(k+1)$, so $byz\in M_{k+1,n-k-1}$ as desired.  Since elements of the form $yz$ span $M_{1,n-k-1}$, we get that $bM_{1,n-k-1}\subset M_{k+1,n-k-1}$.

    Second, let $y\in L(k)$, $z\in \A(n-k)$ with $z$ not in degree $0$, so that $yz\in M_{k,n-k}$.  If $z\not\in\A(n-k-1)$, then $zX_{n-k}$ must be zero in $\A(n-k)$.  Then we may assume $z\in\A(n-k-1)$; we will show that both $yzb'$ and $yzM_{1,n-k-1}$ are contained in $M_{k+1,n-k-1}$.  Now if we write $zb'$ in admissible form, multiplying by $y$ will give an element already in admissible form by looking at degrees, so we need only show that $zb'\in M_{1,n-k-1}$.  If we multiply out using the Adem relations, this is clearly true, modulo showing that $zb'$ contains no terms of length $1$.  This is easy to show; by the Milnor basis, $z$ cannot end in a Steenrod square with degree a multiple of $2^{n-k}$, but $b'$ is a Steenrod square in a degree which is $1$ modulo $2^{n-k}$, so on multiplying the two, the length $1$ term must vanish, as desired.

    %maybe we can merge the following with the previous argument somehow?
    Finally, then, we must show that $yzM_{1,n-k-1}\subset M_{k+1,n-k-1}$; let $y'\in L(1)$ and $z'\in\A(n-k-1)$, so we must show that $yzy'z'\in M_{k+1,n-k-1}$.  Again looking at the Adem relations, we may write $zy'$ in the form $y''z''$ where $y''\in L(1)$ and $z''\in\A(n-k-1)$.  Then $z''z'\in \A(n-k-1)$ and is not in degree zero, and $yy''\in L(k+1)$, so $yy''z''z'\in M_{k+1,n-k-1}$.

    Then the induction is complete; all terms of $tt'-bb'$ are in $M_{k+1,n-k-1}$ as desired.
    \end{proof}

    Then since the elements in question may be written in the given form, their shortest terms are unique and distinct, so the terms are linearly independent.
  \end{proof}

\end{document}

### 2015-06-10 paper.tex
various work on paper

@@
 
-\usepackage{amsmath,amssymb,amsfonts,amsthm}
+\usepackage{amsmath,amssymb,amsfonts,amsthm,graphicx}
 
@@
 \newcommand{\Sq}{\mathrm{Sq}}
+\newcommand{\mmod}{/\!/\!}
+\renewcommand{\L}{\bar{L}}
+\newcommand{\Lkm}[1][k]{L{(#1)}_{-\infty}}
 
@@
 \newtheorem{lem}{Lemma}
+\newtheorem{thm}{Theorem}
+\newtheorem{defn}{Definition}
 
 %TODO: frontmatter
-\title{Determining the Structure of Length-$k$ Steenrod Operations as $A(r)$-Modules}
+\title{Determining the Structure of Length-$k$ Steenrod Operations as $\A(r)$-Modules}
 \author{Benjamin Kraft}
@@
 \begin{abstract}
-  The Steenrod Algebra $\mathcal{A}$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.
-Let $\mathcal{A}(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.
+  The Steenrod Algebra $\A$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.
+Let $\A(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.
 Consider the modules $L(k)$ spanned by sequences of Steenrod operations of length $k$.
-Welcher proved that $L(k)$ is a free module over $\mathcal{A}(k-1)$.
-We are interested in finding the structure of $L(k)$ as an $\mathcal{A}(r)$-module for any $r$.
-We conjecture that $L(k)$ is built as an $\mathcal{A}(r)$-module out of $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, in the sense that it has an increasing filtration with quotients isomorphic to $\mathcal{A}(r)/\!/\!\mathcal{A}(r-k)$, and present partial results towards that claim.
+Welcher proved that $L(k)$ is a free module over $\A(k-1)$.
+We are interested in finding the structure of $L(k)$ as an $\A(r)$-module for any $r$.
+We conjecture that $L(k)$ is built as an $\A(r)$-module out of $\A(r)\mmod\A(r-k)$, in the sense that it has an increasing filtration with quotients isomorphic to $\A(r)\mmod\A(r-k)$, and present partial results towards that claim.
 In addition, we prove some interesting commutation relations in the Steenrod algebra relating to representations of Steenrod Algebra elements in Wood's $Z$-basis.
@@
 
+\section{Introduction}
+
+\begin{defn}
+  Following Welcher~\cite{TODO}, we let $L(k)$ be the subspace of $\A$ spanned by admissible monomials of length $k$.
+\end{defn}
+
+The vector space $L(k)$ may be considered as a (left or right) $\A$-module, by simply doing the multiplication in $\A$, and considering any terms which are not of length $k$ to be zero.
+Similarly, we may consider $L(k)$ as a module over any of the subalgebras $\A(k)$.
+$L(k)$ has a nice periodic structure, which we consider in more detail in the next section.
+
+%TODO: basic results
+
+%TODO: structure of the paper
+
+\section{Definitions and Background}
+
+\subsection{The Periodicity of \boldmath$L(k)$}
+
+\begin{figure}
+  \includegraphics[width=0.9\textwidth]{pics/L2-paper.pdf}
+  \caption{The structure of $\L(2)$ as an $\A(2)$-module.
+Each point is an element of the Adem basis for $\L(2)$; for example the bottommost point in the first column is $\Sq^0$, while the bottommost point in the second column is $\Sq^2\Sq^1$.
+The lines show the action of $\A(2)$; for instance the light gray lines connect each monomial $\Sq^I$ to the terms of $\Sq^4\Sq^I$.
+Shifting the diagram down by 8 rows or right by 4 columns shows the periodic structure.
+$L(2)$ consists of all but the first column; $\Lkm$ extends infinitely far to the left.\label{fig:L2}}
+\end{figure}
+
+If we consider $L(k)$ as a module over $\A(n)$, we find from the Adem relations that it has a nice periodic structure.
+In particular, the left action of $\Sq^{2^i}$ commutes with the operation of decreasing each monomial $\Sq^I$ componentwise by the tuple $(d_1, d_2, \ldots, d_k)$ where $d_j$ is divisible by $2^{i+2-j}$ if $j<i+2$, removing the monomial if it is no longer admissible.
+This fact follows trivially from the Adem relations; in particular adding a large power of two to the exponent of the second term does not affect the binomial coefficients.
+
+In many cases, that periodic structure becomes somewhat easier to consider if we include all admissible monomials of length \emph{at most} $k$.
+\begin{defn}
+  Let $\L(k)=\bigcup_{i\leq k}L(i)$ include all admissible monomials of length \emph{at most} $k$.
+\end{defn}
+
+For example, the structure of $\L(2)$ as an $\A(2)$-module is shown in Figure~\ref{fig:L2}; it is periodic modulo $(8, 4)$.
+
+In some cases it is simpler to dispense with the edge cases of the periodic structure entirely, and simply extend $L(k)$ back to negative infinity (or equivalently, to the left in Figure~\ref{fig:L2}).
+Formally, we let $\Lkm$ consist of admissible monomials composed of factors $\Sq^i$, where $i$ may be negative.
+(Admissible monomials are still defined as those in which each term is at least twice the previous one, so that $\Sq^3\Sq^{-2}$ and $\Sq^3\Sq^{-1}$ are both admissible, but $\Sq^{-3}\Sq^{-1}$ is not.)  The module structure of $\Lkm$ over $\A(k)$ may be defined by extending the periodic structure of $L(k)$ backwards, so that $\Sq^{2^i}\Sq^I$ is defined by adding some sufficiently large admissible tuple $J=(2^{i+1}m_1, 2^im_2, \ldots)$ to $I$ so that it becomes positive, performing the multiplication, and then subtracting the tuple $J$ from the exponent of each term of the result, discarding any terms which are no longer admissible.
+
+\subsection{The Wood \boldmath$Z$ basis}
+In addition to the Adem basis, the $Z$ basis described by Wood~\cite{TODO} will be important to our work.  Let
+\[ X_n = \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}, \]
+and let $Z_n = X_n X_{n-1}\cdots X_0$.  Then Wood showed that $Z_n$ is the top element of $\A(n)$, and furthermore that the following is a basis for $\A(n)$.
+\begin{defn}
+  The Wood $Z$ basis for $\A(n)$ consists of the set of monomials in Steenrod squares produced by multiplying any subset of the Steenrod squares in $Z_n$ in order.
+\end{defn}
+
+The Steenrod Algebra, then, consists of all finite subproducts of the infinite product
+\[ \cdots\Sq^8\Sq^{12}\Sq^{14}\Sq^{15}\Sq^4\Sq^6\Sq^7\Sq^2\Sq^3\Sq^1.\]
+
+For example,
+\[Z_2 = \Sq^4\Sq^6\Sq^7\Sq^2\Sq^3\Sq^1 = \Sq^17 \Sq^5 \Sq^1\]
+is the top class of $\A(2)$.  Furthermore, the following set forms a basis of $\A(2)$:
+\begin{gather*}
+  \Sq^4\Sq^6\Sq^7\Sq^2\Sq^3\Sq^1 \\
+  \Sq^4\Sq^6\Sq^7\Sq^2\Sq^3 \quad \Sq^4\Sq^6\Sq^7\Sq^2\Sq^1 \quad \Sq^4\Sq^6\Sq^7\Sq^3\Sq^1 \\
+  \Sq^4\Sq^6\Sq^2\Sq^3\Sq^1 \quad \Sq^4\Sq^7\Sq^2\Sq^3\Sq^1 \quad \Sq^6\Sq^7\Sq^2\Sq^3\Sq^1 \\
+  \vdots \\
+  \Sq^4 \quad \Sq^6 \quad \Sq^7 \quad \Sq^2 \quad \Sq^3 \quad \Sq^1 \\
+  1
+\end{gather*}
+
+Note that the $Z$ basis builds larger $\A(n)$ from right to left: in particular, $Z_n = X_n Z_{n-1}$, and therefore each element of the $Z$ basis for $\A(n)$ is formed by multiplying some subproduct of $X_n$ by an element of the $Z$ basis for $\A(n-1)$.
+This means that the algebras $\A(n)\mmod\A(n-k)$ have a particularly nice representation in the $Z$-basis.
+
+\begin{prop}
+  The set of monomials produced by multiplying any subset of the Steenrod squares in $X_n X_{n-1}\cdots X_{n-k+1}$ in any order forms a set of representatives for a basis for the algebra $\A(n)\mmod\A(n-k)$.
+In particular, $X_n X_{n-1} \cdots X_{n-k+1}$ is a representative of its top class.
+\end{prop}
+\begin{proof}
+  Consider some element $z$ of the $Z$ basis for $\A(n)$, which must be a subproduct of $X_n X_{n-1}\cdots X_{n-k+1}Z_{n-k}$.
+If it contains any Steenrod squares from $Z_{n-k}$, then it can be written $z = xy$ where $x\in \A(n)$ and $y\in \A(n-k)$.
+Then in $\A(n)\mmod \A(n-k) = \A(n) \otimes_{\A(n-1)} \F_2$, $z$ is equal to zero.
+Otherwise, it is in the proposed basis.
+So since a basis of $\A(n)$ must span $\A(n) \mmod \A(n-k)$, the proposed basis must span.
+Furthermore, $\A(n)$ has dimension $2^{2^{n+1}-1}$, and therefore $\A(n)\mmod \A(n-k)$ has dimension $2^{2^{n+1}-2^{n-k+1}}$, so since $X_n X_{n-1} \cdots X_{n-k+1}$ is a product of $2^{n+1}-2^{n-k+1}$ Steenrod squares, there are the right number of subproducts, so they must form a basis.
+The element $X_n X_{n-1} \cdots X_{n-k+1}$ is the only element in the right degree to be a representative of the top class, so it must in fact be.
+\end{proof}
+
+For example,
+\[ \Sq^4\Sq^6\Sq^7\Sq^2\Sq^3 \]
+is a representative of the top class of $\A(2)\mmod \A(0)$, and the following set forms a basis of $\A(2)\mmod \A(0)$:
+\begin{gather*}
+  \Sq^4\Sq^6\Sq^7\Sq^2\Sq^3 \\
+  \Sq^4\Sq^6\Sq^7\Sq^2 \quad \Sq^4\Sq^6\Sq^7\Sq^3 \\
+  \Sq^4\Sq^6\Sq^2\Sq^3 \quad \Sq^4\Sq^7\Sq^2\Sq^3 \quad \Sq^6\Sq^7\Sq^2\Sq^3 \\
+  \vdots \\
+  \Sq^4 \quad \Sq^6 \quad \Sq^7 \quad \Sq^2 \quad \Sq^3 \\
+  1
+\end{gather*}
+
 \section{The Commutation Relation}
 
-  Recall that, in the Wood Z-basis, the top class of $A(n)\otimes_{A(n-1)}\F_2$ may be written as $\Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}$; denote this by $X_n$.
+We now prove a commutation relation involving the $X_n$ which is interesting in its own right, and will prove useful in the remainder of the paper.
 
@@
     The Adem relations give us that
-      \[\Sq^{2^n-1}\Sq^{2^{n+1}m} = \sum_{j=0}^{2^{n-1}-1}\binom{2^{n+1}m-1-j}{2^n-1-2j}\Sq^{2^{n+1}m+2^n-1-j}\Sq^j.\]
+  \begin{equation}\Sq^{2^n-1}\Sq^{2^{n+1}m} = \sum_{j=0}^{2^{n-1}-1}\binom{2^{n+1}m-1-j}{2^n-1-2j}\Sq^{2^{n+1}m+2^n-1-j}\Sq^j.\label{eq:identity-1-adem}\end{equation}
     Recall that in $\F_2$, $\binom{p}{q}$ is $1$ if and only if $p$ has a one in every position of its binary expansion that $q$ has a $1$.
 Let $2^n>j>0$ and let $2^{\omega(j)}$ be the largest power of $2$ dividing $j$.
-Then the binary expansion of $2^{n+1}m-1-j$ ends in a zero and then $\omega(j)$ ones, while that of $2^n-1-2j$ ends in a zero and then $\omega(j)+1$ ones, so the binomial coefficient in the Adem relation is zero.
-Then the only term of the Adem relation that is nonzero is that when $j=0$, where $2^{n+1}m-1$ ends with $n+1$ ones, while $2^n-1$ consists of $n$ ones.
+Then the binary expansion of $2^{n+1}m-1-j$ ends in $\omega(j)$ ones preceded by a zero, while that of $2^n-1-2j$ ends in a zero and then $\omega(j)+1$ ones, so the binomial coefficient in the Adem relation is zero.
+Then the only term of the Adem relation in~\eqref{eq:identity-1-adem} that is nonzero is that when $j=0$, where $2^{n+1}m-1$ ends with $n+1$ ones, while $2^n-1$ consists of $n$ ones.
+The result follows immediately.
   \end{proof}
@@
 
-      Now let $2^{n-1}-2^{k-1}<j<2^n-2^{k-1}$; we will show that the binomial coefficient
+    Now let $2^{n-1}-2^{k-1}<j<2^n-2^{k-1}$; we will show that the latter binomial coefficient
     \[\binom{2^{n+1}m-2^{k-1}-1-j}{2^{n+1}-2^k-2j}\]
@@
 
-      Finally, if $j=2^n-2^{k-1}$, the binomial coefficient in the second formula has a zero on the bottom, thus is one, so we get the middle term of the claimed identity.
+    Finally, if $j=2^n-2^{k-1}$, the binomial coefficient in the second formula has a zero on the bottom, thus is one, so we get $\Sq^{2^n-2^k+2^{n+1}m}\Sq^{2^n-2^{k-1}}$, the middle term of the claimed identity, as the only leftover term, and the identity holds.
   \end{proof}
@@
 
-\section{Representatives for \boldmath$\A(n)//\A(n-k)$}
+\section{Representatives for \boldmath$\A(n)\mmod\A(n-k)$}
 
-  For convenience let $\bar{L}(k)=\bigcup_{i\leq k}L(i)$.
+\begin{prop}\label{prop:set-of-reps}
+  % There is a set of representatives for a basis of $\A(n)\mmod\A(n-k)$ (in the sense of the above note), all of which lie in $\L(k)$.
+  Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \L(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.
+Equivalently, $\A(n)\subset \L(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
+\end{prop}
 
-  \begin{prop}
-    % There is a set of representatives for a basis of $\A(n)//\A(n-k)$ (in the sense of the above note), all of which lie in $\bar{L}(k)$.
-    Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \bar{L}(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.
-Equivalently, $\A(n)\subset \bar{L}(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
-  \end{prop}
+Proposition~\ref{prop:set-of-reps} essentially says that there exists a set of representatives for a basis of $\A(n)\mmod\A(n-k)$, all of which lie in $\L(k)$.
 
@@
   We induct on the number of squares selected.
-The empty product is $1$, which is in $\bar{L}(k)$, so we are done.
+The empty product is $1$, which is in $\L(k)$, so we are done.
 So now suppose the proposition holds on all elements of the Wood Z-basis of length at most $l-1$, and let $t$ be an element of length $l$.
@@
   Now $\Sq^myz$ is clearly of the prescribed form, since $\Sq^my$ is in $\A$.
-If $x\in\bar{L}(k-1)$, then $\Sq^m x\in \bar{L}(k)$ (since application of the Adem relations cannot increase the length of a monomial), so we may suppose $x\in L(k)$.
+If $x\in\L(k-1)$, then $\Sq^m x\in \L(k)$ (since application of the Adem relations cannot increase the length of a monomial), so we may suppose $x\in L(k)$.
 
@@
 
-  \begin{lem}\label{lem:linearly-independent}
-    Taken as elements of $L(k)_{-\infty}$ as a module over $\A(n)\otimes_{\A(n-k)}\F_2$, the set
-    \[\{X_nX_{n-1}\cdots X_{n-k+1}\Sq^I\mid I=(2^{n+1}m_1,2^nm_2,\ldots,2^{n-k+2}m_k), m_1\geq m_2\geq\cdots\geq m_{n-k}\}\]
+\begin{lem}\label{lem:li-vs}
+  %TODO: is the claim here really that they are LI in the vector space over $\F2$?
+  Taken as elements of $\Lkm$ as a vector space over $\F_2$, the set
+  \[\{X_nX_{n-1}\cdots X_{n-k+1}\Sq^I\mid I=(2^{n+1}m_1,2^nm_2,\ldots,2^{n-k+2}m_k), m_1\geq\cdots\geq m_k\}\]
   is linearly independent.
@@
   \[X_n\Sq^{2^{n+1}m_1}X_{n-1}\Sq^{2^nm_2}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_k}.\]
-    Let $M_{k,n}$ be the subspace of $\A$ which is spanned by elements of the form $yz$ where $y\in L(k)$ and $z\in \A(n)$ with the degree of $z$ at least $1$.
+  Let $M_{k,n}$ be the subspace of $\A$ which is spanned by monomials of the form $yz$ where $y\in L(k)$ and $z\in \A(n)$ with the degree of $z$ at least $1$.
 If $b\in \A$, we will write $b+M_{k,n}$ for the subset of $\A$ consisting of elements of the form $b+m$ for $m\in M_{k,n}$.
@@
 
+    %TODO: fix gap here: why isn't $X_n$ just in $M_{1, n-1}$ (may just need to show this out)?
   We proceed by induction on $k$ for fixed $n$.
@@
   Now suppose the proposition holds for some $k$; we will show it for $k+1$.  Let 
-    \[t=X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}},\]
-    let $t'=X_{n-k}\Sq^{2^{n-k+1}m_{n-k}}$, let
-    \[b=\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1},\]
-    and let $b'=\Sq^{2^{n-k+1}(m_{n-k}+n-k)+1}$.
+  \begin{align*}
+  t&=X_n\Sq^{2^{n+1}m_n}X_{n-1}\Sq^{2^nm_{n-1}}\cdots X_{n-k+1}\Sq^{2^{n-k+2}m_{n-k+1}} \\
+  t'&=X_{n-k}\Sq^{2^{n-k+1}m_{n-k}} \\
+  b&=\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1} \\
+  b'&=\Sq^{2^{n-k+1}(m_{n-k}+n-k)+1}.
+  \end{align*}
   Then by inductive hypothesis we know that $t\in b+M_{k,n-k}$ and by the base case we know that $t'\in b'+M_{1,n-k-1}$; we wish to show that $tt'\in bb'+M_{k+1,n-k-1}$.
@@
 
+  %TODO: check this more carefully.
   First, let $y\in L(1)$, $z\in \A(n-k-1)$ with $z$ not in degree $0$, so that $yz\in M_{1,n-k-1}$.
@@
   Second, let $y\in L(k)$, $z\in \A(n-k)$ with $z$ not in degree $0$, so that $yz\in M_{k,n-k}$.
-If $z\not\in\A(n-k-1)$, then $zX_{n-k}$ must be zero in $\A(n-k)$.
+If $z\not\in\A(n-k-1)$, then $zX_{n-k}$ and thus $yzt'$ must be zero in $\A(n-k)$.
 Then we may assume $z\in\A(n-k-1)$; we will show that both $yzb'$ and $yzM_{1,n-k-1}$ are contained in $M_{k+1,n-k-1}$.
@@
 
+  %TODO: show somewhere that the shortest terms are in fact the shortest, i.e., that all terms of $M_{k, n-k}$ are of length at least $k+1$.
+  %TODO: make this argument clearer
   Then since the elements in question may be written in the given form, their shortest terms are unique and distinct, so the terms are linearly independent.
@@
 
+\section{Filtrations of \boldmath$L(k)$}
+
+\subsection{The General Case}
+
+%TODO: I think we can't use the word basis here, we have to talk about filtrations, or "bases".
+Lemma~\ref{lem:li-vs} suggests that the set
+\[ B_{n,k} = \{Sq^I \mid I = (2^{n+1}m_1, 2^n m_2, \ldots, 2^{n-k+2}m_k), m_1 \geq \cdots \geq m_k \} \]
+may be a basis for $\Lkm$ as an $\A(n)\mmod\A(n-k)$-module, through the following theorem.
+
+\begin{thm}\label{thm:li-module}
+  The set $B_{n,k}$ is linearly independent as elements of $\Lkm$ considered as an $\A(n)\mmod\A(n-k)$-module.
+\end{thm}
+
+\begin{proof}
+  Suppose it were not: suppose that there were some relation $\sum_i x_i b_i = 0$, with $x_i\in\A(n)\mmod\A(n-k)$ and $b_i\in B_{n,k}$.
+Let $z$ be the top class of $\A(n)\mmod\A(n-k)$.
+We now have two cases.
+
+  %TODO: check that hopf algebra is the right criterion here, and that it's true
+  First, consider the case where not all $b_i$ are in the same degree; let $b_j$ be one of the ones in the highest possible degree.
+Since $\A(n)\mmod\A(n-k)$ is a Hopf algebra, there exists some $y_j$ in $\A(n)\mmod\A(n-k)$ such that $y_j x_j = z$.
+Then for any $b_{j'}$ in a lower degree than $b_j$, the corresponding $x_{j'}$ will be in a higher degree than $x_j$.
+Then $y_j x_j = 0$, so multiplying the entire relation by $y_j$ leaves a new relation amongst only the $b_i$ in the highest possible degree.
+
+  So now we consider the case where all $b_i$ are in the same degree.
+As before, let $y$ be such that $y x_1 = z$.
+Then in the relation $\sum_i y x_i b_i = 0$, the first coefficient is $z$, and all of the later ones are either $z$ or zero.
+But each $z b_i$ is an element of the linearly independent set from Lemma~\ref{lem:li-vs}, so there can be no relation among them.
+Then we have a contradiction: there is no relation among the $b_i$ over $\A(n)\mmod\A(n-k)$, so the set $B_{n,k}$ is linearly independent.
+\end{proof}
+
+Given this, it remains only to show that in some sense the basis $B_{n,k}$ has the right size.
+Unfortunately, this is more difficult than it sounds.
+If we take the full basis $B_{n,k}$ over $\Lkm$, the vector space of elements in each degree is infinite-dimensional, so the argument cannot proceed.
+If we consider only the intersection of $B_{n,k}$ with $\L(k)$, this difficulty is removed, but another one presents itself: some classes are not spanned by the $B_{n,k}$.
+While this issue is irrelevant to the desired final result, in that the periodic structure is still entirely encapsulated, with the exception of some initial conditions, a dimension-counting argument would require a careful accounting of these initial exceptions, which we found intractible in the general case.
+
+Thus we were unable to show the desired result in the general case.
+We do believe that a proof could proceed along these lines, we have just been unable to complete the dimension-counting portion of the argument, and hope that future research will yield fruits in the area.
+
+\subsection{The case of small \boldmath$n$ and \boldmath$k$}
+
+%TODO: discussion of small cases
+
+%TODO: future work
+
 \end{document}

### 2015-06-11 paper.tex
more writing

@@
 
+%TODO: more intro
+
 \begin{defn}
@@
 
-%TODO: basic results
+%TODO: future work
 
@@
 \section{Definitions and Background}
+
+\subsection{The Adem Relations}
+
+Recall that the Steenrod squares satisfy the Adem relations:
+\[ \Sq^i\Sq^j = \sum_{m=0}^{m/2}\binom{j-m-1}{i-2m}\Sq^{i+j-k}\Sq^k .\]
+The binomial coefficient is taken modulo 2, and it is a standard result in combinatorics that it is nonzero if and only if for each bit of the binary representation of $i-2m$ that is 1, the corresponding bit of $j-m-1$ must also be 1.
+We will use this fact extensively.
 
@@
 
-\begin{prop}\label{prop:set-of-reps}
+\begin{lem}\label{lem:set-of-reps}
   % There is a set of representatives for a basis of $\A(n)\mmod\A(n-k)$ (in the sense of the above note), all of which lie in $\L(k)$.
   Any element $t$ of $\A(n)$ may be written as $x + yz$ where $x\in \L(k)$, $y\in \A$, and $z\in\A(n-k)\setminus \F_2$ for any $k\geq 0$.
-Equivalently, $\A(n)\subset \L(k) + \A(\Sq^i)_{i\leq k}$, where we take $\A(\Sq^i)_{i\leq k}$ to be the left ideal of $\A$ generated by $\{\Sq^i\}_{i\leq k}$, for any $k\geq 0$.
-\end{prop}
+Equivalently, $\A(n)\subset \L(k) + \A(\Sq^{2^i})_{i\leq n-k}$, where we take $\A(\Sq^{2^i})_{i\leq n-k}$ to be the left ideal of $\A$ generated by $\{\Sq^{2^i}\}_{i\leq n-k}$, for any $k\geq 0$.
+\end{lem}
 
-Proposition~\ref{prop:set-of-reps} essentially says that there exists a set of representatives for a basis of $\A(n)\mmod\A(n-k)$, all of which lie in $\L(k)$.
+Proposition~\ref{lem:set-of-reps} essentially says that there exists a set of representatives for a basis of $\A(n)\mmod\A(n-k)$, all of which lie in $\L(k)$.
 
 \begin{proof}
-  It suffices to prove the proposition on some basis for $\A(n)$.
-Recall that in Wood's Z-basis of the Steenrod algebra, a basis for $\A(n)$ consists of all strings of squares selected from~\eqref{eq:woodz} and multiplied in the given order:
-  \begin{equation} \label{eq:woodz}
-    \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1} \Sq^{2^n-2^{n-1}}\cdots\Sq^{2^n-1}\cdots\Sq^2\Sq^3\Sq^1
-  \end{equation}
-
-  We induct on the number of squares selected.
+  It suffices to prove the proposition on some basis for $\A(n)$; we use the Wood $Z$ basis.
+We induct on the number of squares selected to form the basis element.
 The empty product is $1$, which is in $\L(k)$, so we are done.
@@
   Write $x$ in admissible form as $\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$.
-If $\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$ is already in admissible form, then we have that $2^ki_k\leq 2^{k-1}i_{k-1}\leq\cdots\leq2i_1\leq m<2^{n+1}$, so in particular, $i_k<2^{n-k+1}$, so $i_k\in\A(\Sq^i)_{i\leq k}$, so in fact the entire product $\Sq^mx$ is in $\A(\Sq^i)_{i\leq k}$.
+If $\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k}$ is already in admissible form, then we have that $2^ki_k\leq 2^{k-1}i_{k-1}\leq\cdots\leq2i_1\leq m<2^{n+1}$, so in particular, $i_k<2^{n-k+1}$, so $\Sq^{i_k}\in\A(\Sq^{2^i})_{i\leq n-k}$, so in fact the entire product $\Sq^mx$ is in $\A(\Sq^{2^i})_{i\leq n-k}$.
 Then we may apply the Adem relations at least once, to get
-  \[\Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k} = a_0\Sq^{m+i_1}\Sq^{i_2}\cdots\Sq^{i_k} + \sum_{1\leq j\leq \frac{m}2}a_j\Sq^{m+i_1-j}\Sq^j\Sq^{i_2}\cdots\Sq^{i_k};\]
+  \begin{multline*}
+  \Sq^m\Sq^{i_1}\Sq^{i_2}\cdots\Sq^{i_k} = a_0\Sq^{m+i_1}\Sq^{i_2}\cdots\Sq^{i_k} \\
+                  + \sum_{1\leq j\leq \frac{m}2}a_j\Sq^{m+i_1-j}\Sq^j\Sq^{i_2}\cdots\Sq^{i_k};
+  \end{multline*}
   the first term we know is in $L(k)$.
-For each of the remaining terms, if we may repeat the same process, ignoring the first factor; either the product is in $\A(\Sq^i)_{i\leq k}$, or it may be further reduced by the Adem relations.
-Finally, if we apply the Adem relations $k$ times, the rightmost factor must be $\Sq^i$ for $i\leq \frac{m}{2^k}<2^{n-k+1}$, so similarly the product lies in the ideal $\A(\Sq^i)_{i\leq k}$.
+For each of the remaining terms, if we may repeat the same process, ignoring the first factor; either the product is in $\A(\Sq^{2^i})_{i\leq n-k}$, or it may be further reduced by the Adem relations.
+Finally, if we apply the Adem relations $k$ times, the rightmost factor must be $\Sq^i$ for $i\leq \frac{m}{2^k}<2^{n-k+1}$, so similarly the product lies in the ideal $\A(\Sq^{2^i})_{i\leq n-k}$.
 Then $\Sq^mx$ may be written in the desired form, so we are done.
@@
 \begin{lem}\label{lem:li-vs}
-  %TODO: is the claim here really that they are LI in the vector space over $\F2$?
   Taken as elements of $\Lkm$ as a vector space over $\F_2$, the set
@@
 
-    %TODO: fix gap here: why isn't $X_n$ just in $M_{1, n-1}$ (may just need to show this out)?
     We proceed by induction on $k$ for fixed $n$.
-If $k=1$, we know that $X_n\in M_{1,n-1}+\Sq^{n2^{n+1}+1}$, since the latter term is the only element of $L(1)$ in the right dimension.
+If $k=1$, we know by Lemma~\ref{lem:set-of-reps} that $X_n\in M_{1,n-1}+\Sq^{n2^{n+1}+1}$; the latter term is the only element of $\L(1)$ in the right dimension, so we know it must be the value of $x$ from the lemma.
 Then since, as an $\A(n)$-module, $L(1)$ is periodic in degrees modulo $2^{n+1}$, $X_n\Sq^{2^{n+1}m}\in M_{1,n-1}+\Sq^{2^{n+1}(n+m)+1}$.
@@
   %TODO: show somewhere that the shortest terms are in fact the shortest, i.e., that all terms of $M_{k, n-k}$ are of length at least $k+1$.
-  %TODO: make this argument clearer
-  Then since the elements in question may be written in the given form, their shortest terms are unique and distinct, so the terms are linearly independent.
+  Note that the terms 
+\begin{equation}\Sq^{2^{n+1}(m_n+n)+1}\Sq^{2^n(m_{n-1}+n-1)+1}\cdots\Sq^{2^{n-k+2}(m_{n-k+1}+n-k+1)+1}\label{eq:b-term}\end{equation}
+  are outside of $M_{k, n-k}$ since all elements of $\A(n-k)$ have degree smaller than the last factor of~\eqref{eq:b-term}.
+Then since the elements in question may be written in the given form, we know that each has a unique term outside of $M_{k, n-k}$.
+Those terms are distinct, so the terms are linearly independent.
 \end{proof}
@@
 
-%TODO: I think we can't use the word basis here, we have to talk about filtrations, or "bases".
 Lemma~\ref{lem:li-vs} suggests that the set
 \[ B_{n,k} = \{Sq^I \mid I = (2^{n+1}m_1, 2^n m_2, \ldots, 2^{n-k+2}m_k), m_1 \geq \cdots \geq m_k \} \]
-may be a basis for $\Lkm$ as an $\A(n)\mmod\A(n-k)$-module, through the following theorem.
+may allow us to construct a filtration of $\Lkm$ with copies of $\A(n)\mmod\A(n-k)$ as the filtration quotients, through the following theorem.
 
 \begin{thm}\label{thm:li-module}
-  The set $B_{n,k}$ is linearly independent as elements of $\Lkm$ considered as an $\A(n)\mmod\A(n-k)$-module.
+  The set $B_{n,k}$ is linearly independent in $\Lkm$ with respect to coefficients taken from $\A(n)\mmod\A(n-k)$.
 \end{thm}
@@
 
-  %TODO: check that hopf algebra is the right criterion here, and that it's true
   First, consider the case where not all $b_i$ are in the same degree; let $b_j$ be one of the ones in the highest possible degree.
-Since $\A(n)\mmod\A(n-k)$ is a Hopf algebra, there exists some $y_j$ in $\A(n)\mmod\A(n-k)$ such that $y_j x_j = z$.
+Since $\A(n)\mmod\A(n-k)$ is a Poincar\'e algebra, there exists some $y_j$ in $\A(n)\mmod\A(n-k)$ such that $y_j x_j = z$.
 Then for any $b_{j'}$ in a lower degree than $b_j$, the corresponding $x_{j'}$ will be in a higher degree than $x_j$.
@@
 
-Given this, it remains only to show that in some sense the basis $B_{n,k}$ has the right size.
+Given this, it remains only to show that in some sense the ``basis'' $B_{n,k}$ has the right size.
 Unfortunately, this is more difficult than it sounds.
-If we take the full basis $B_{n,k}$ over $\Lkm$, the vector space of elements in each degree is infinite-dimensional, so the argument cannot proceed.
+If we take the full $B_{n,k}$, the vector space of elements in each degree of $\Lkm$ is infinite-dimensional, so the argument cannot proceed.
 If we consider only the intersection of $B_{n,k}$ with $\L(k)$, this difficulty is removed, but another one presents itself: some classes are not spanned by the $B_{n,k}$.
-While this issue is irrelevant to the desired final result, in that the periodic structure is still entirely encapsulated, with the exception of some initial conditions, a dimension-counting argument would require a careful accounting of these initial exceptions, which we found intractible in the general case.
+While these are irrelevant to the desired final result, in that the periodic structure is still entirely encapsulated, with the exception of some initial conditions, a dimension-counting argument would require a careful accounting of these initial exceptions, which we found intractible in the general case.
 
@@
 
-%TODO: discussion of small cases
+Fortunately, in small cases the dimension-counting is tractable.
+We begin by considering $\Lkm[1]$ as a module over some $\A(n)$.
+The space $\Lkm[1]$ is particularly simple, since it is one-dimensional in each integral degree.
 
-%TODO: future work
+\begin{thm}
+  As an $\A(n)$-module, $\Lkm[1]$ has a filtration with each filtration quotient consisting of a copy of $\A(n)\mmod\A(n-1)$.
+\end{thm}
+\begin{proof}
+  We begin by showing that $\L(1)$ has such a filtration, with a single finite-dimensional base space of the filtration not isomorphic to $\A(n)\mmod\A(n-1)$.
+The filtration quotients, in particular, are each a copy of $\A(n)\mmod\A(n-1)$ built on each element of $B_{n,1}\cap\L(1)$.
+
+  These quotients are linearly independent by Theorem~\ref{thm:li-module}.
+In addition, I claim they contain one class in each degree above $(n-1)2^{n+1}$.
+In particular, consider some degree $d = m2^{n+1} - d'$, where $d' < 2^{n+1}$ and $m \geq n$.
+Suppose that the binary representation of $d'$ has ones in the bits $k_1, k_2, \ldots, k_j$.
+Then the class
+  \[\Sq^{2^{n+1}-2^{k_1}}\Sq^{2^{n+1}-2^{k_2}}\cdots\Sq^{2^{n+1}-2^{k_j}}\Sq^{(m-j)2^{n+1}} \]
+  has is in the correct degree and is in the proposed filtration.
+So the proposed filtration indeed spans all of $\L(1)$ with all quotients copies of $\A(n)\mmod\A(n-1)$ except possibly for some finite-dimensional base space containing classes in small degrees.
+
+  Since the structure of $\Lkm[1]$ is periodic, this finite-dimensional base disappears if we extend the filtration to $\Lkm[1]$, so it has a filtration with all quotients consisting of a copy of $\A(n)\mmod\A(n-1)$ as desired.
+\end{proof}
+
+%TODO: pictures of the filtrations of L(1) and L(2)?
+
+Next we consider the case of $\L(2)$.
+Welcher's result shows that $\L(2)$ is free as an $\A(1)$-module.
+As an $\A(2)$-module, we wish to show that $\Lkm[2]$ admits a filtration with quotients $\A(2)\mmod\A(0)$.
+Again, only the dimension-counting remains.
+
+\begin{thm}
+  As an $\A(2)$-module, $\Lkm[2]$ has a filtration with each filtration quotient consisting of a copy of $\A(n)\mmod\A(n-1)$.
+\end{thm}
+\begin{proof}
+  Again we begin by showing that $\L(2)$ has a filtration, this time with a single base space which we construct, and which is finite in each degree, and each other filtration quotient is isomorphic to $\A(2)\mmod\A(0)$, again built on elements of $B_{2,2}\cap \L(2)$.
+
+  These quotients are linearly independent by Theorem~\ref{thm:li-module}.
+In addition, I claim that none of the filtration quotients built on this basis span any of the following basis elements of $\L(2)$:
+\begin{equation} \Sq^1, \Sq^3\Sq^1, \Sq^4\Sq^1, \Sq^5\Sq^1, \ldots. \label{eq:extra}\end{equation}
+  This is easy to see from the Adem relations: clearly the only elements from $B_{2,2}\cap \L(2)$ that could be multiplied by elements of $\A(2)$ to get anything ending in $\Sq^1$ are the $\Sq^{8i}$, and it is easy to see from the Adem relations that any product which could create a term in~\eqref{eq:extra} and no others must involve multiplying by something ending in $\Sq^1$, i.e., by something which is zero in $\A(2)\mmod\A(0)$.
+(Note that this is the point where using $\L(2)$ rather than $L(2)$ makes things much simpler: the set of excluded elements would be much more complicated in $L(2)$.)
+
+  We now use Poincar\'e series to show that after adding those elements from~\eqref{eq:extra}, the claimed filtration quotients add up to have the right dimensionality in each degree, showing that the claimed filtration exists.
+Recall that the Poincar\'e series of a graded vector space is the infinite series $v(x) = \sum_{i=0}^{\infty} x^i \dim V_i$.
+By definition, the Poincar\'e series of $\L(2)$ is
+  \[ l(x) = 1 + x + x^2 + 2x^3 + 2x^4 + 2x^5 + 3x^6 + \cdots = (1 + x + x^2 + \cdots)(1 + x^3 + x^6 + \cdots).\]
+  Furthermore, from looking at the $Z$ basis, we can see that the Poincar\'e series of $\A(2)\mmod\A(0)$ is
+  \[ a(x) = (1+x^4)(1+x^6)(1+x^7)(1+x^2)(1+x^3).\]
+  The Poincar\'e series of the set $B_{2,2}\cap \L(2)$ is
+  \[ b(x) = (1 + x^8 + x^{16} + \cdots)(1 + x^{12} + x^{24} + \cdots).\]
+  Finally, the Poincar\'e series of the extra elements from~\eqref{eq:extra} is
+  \[ c(x) = x + x^3 + x^4 + x^5 + \cdots.\]
+  We wish to show that $l(x) = a(x)b(x) + c(x)$.  This is just a matter of algebra:
+  \begin{align*}
+    a(x)b(x) + c(x) &= (1+x^7)(1 + x^2 + x^4 + \cdots)(1 + x^3 + x^6 + \cdots) + c(x) \\
+                    &= (1-x+x^2-\cdots+x^6)(1 + x + x^2 + \cdots)(1 + x^3 + x^6 + \cdots) \\
+                    &\qquad + x + x^3(1 + x + x^2 + \cdots) \\
+                    &= (1 + x + x^2 + \cdots)(1 - x + x^2 + x^3 + x^6 + x^9 + \cdots) + x \\
+                    &= l(x) + x(x-1)(1 + x + x^2 + \cdots) + x \\
+                    &= l(x) - x + x \\
+    a(x)b(x) + c(x) &= l(x)
+  \end{align*}
+  Then since this filtration has linearly independent filtration quotients, and has the right total dimensionality, it must span the entirety of $\L(2)$.
+As we continue backwards to $\Lkm[2]$, the excluded elements are not a part of the periodic structure, so they get filled in, and we get that $\Lkm[2]$ has a filtration consisting only of copies of $\A(2)\mmod\A(0)$.
+\end{proof}
+
+%TODO: acks
+
+%TODO: bib
 

### 2015-06-11 paper.tex
Add bibliography, and other stuff

@@
 \usepackage{amsmath,amssymb,amsfonts,amsthm,graphicx}
+
+\usepackage[maxnames=99,isbn=false,eprint=true,url=false]{biblatex}
+\addbibresource{paper.bib}
+\renewbibmacro{in:}{}
 
@@
 
-%TODO: frontmatter
 \title{Determining the Structure of Length-$k$ Steenrod Operations as $\A(r)$-Modules}
@@
 
-%TODO: more intro
+The structure of the Steenrod Algebra $\A$ and its various subalgebras has been the object of much study.
+Consider in particular the length filtration of $\A$, with filtration quotients $L(k)$ spanned by sequences of Steenrod operations of length $k$.
+
+%TODO: add more references for context
+\begin{defn}
+  Following Welcher~\cite{welcher}, we let $L(k)$ be the subspace of $\A$ spanned by admissible monomials of length $k$.
+\end{defn}
+
+The vector space $L(k)$ may be considered as a (left) $\A$-module, by simply doing the multiplication in $\A$, and considering any terms which are not of length $k$ to be zero.
+Similarly, we may consider $L(k)$ as a module over any of the subalgebras $\A(k)$.
+$L(k)$ has a nice periodic structure, which we consider in more detail in the next section.
+Welcher~\cite{welcher} proved that $L(k)$ is a free module over the subalgebra $\A(k-1)$ generated by $\Sq^1, \Sq^2, \ldots, \Sq^{2^{k-1}}$.
+We consider $L(k)$ as a module over some larger $\A(n)$, and wish to determine its stucture.
 
 \begin{defn}
-  Following Welcher~\cite{TODO}, we let $L(k)$ be the subspace of $\A$ spanned by admissible monomials of length $k$.
+  Let $\A(n)\mmod\A(n-k)$ be the algebra $\A(n)\otimes_{\A(n-k)}\F_2$, where we consider $\F_2$ as an $\A(n-k)$ module by letting the action of all elements outside of $\F_2$ be zero.
 \end{defn}
 
-The vector space $L(k)$ may be considered as a (left or right) $\A$-module, by simply doing the multiplication in $\A$, and considering any terms which are not of length $k$ to be zero.
-Similarly, we may consider $L(k)$ as a module over any of the subalgebras $\A(k)$.
-$L(k)$ has a nice periodic structure, which we consider in more detail in the next section.
+%TODO: pictures of small ones
+In particular, $\A(n)\mmod\A(n-k)$ has as representatives all elements of $\A(n)$ which cannot be formed by multiplying an element of $\A(n)$ by a nontrivial element of $\A(n-k)$.
+We conjecture that $L(k)$, when considered as an $\A(n)$-module, is built up out of $\A(n)\mmod\A(n-k)$, in the sense that it admits a filtration which, except for some aperiodic parts at the beginning, has each filtration quotient isomorphic to $\A(n)\mmod\A(n-k)$.
 
-%TODO: future work
+Unfortunately, we were unable to show this result in the general case.
+However, we show a part of the result: that there exists a basis for such a filtration which is linearly independent.
+It remains only to show that the basis has the right dimensionality.
+This is more difficult than it is in the case Welcher considers, since we need to keep careful control over the aperiodic parts at the beginning.
 
-%TODO: structure of the paper
+In a few small cases, we showed that the desired result holds: the case of $L(1)$ as an $\A(n)$-module, and the case of $L(2)$ as an $\A(2)$-module.
+We expect that with additional algebra, possibly on a computer, the same method can be applied to any fixed small case, and hope that a similar method can be extended to prove the general case.
 
-\section{Definitions and Background}
+Finally, we show and interesting commutation relation among elements of the Steenrod algebra: if we let
+\[ X_n = \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1} \]
+be defined as in Wood~\cite{wood}, then for any positive integers $m$ and $n$, 
+\begin{equation}X_nX_{n-1}\Sq^{2^{n+1}m} = X_n\Sq^{2^{n+1}m}X_{n-1}.\label{eq:commutation-relation}\end{equation}
+This relation is useful in proving our linear independence result, and we hope it can be useful in general.
+
+In Section~\ref{sec:defns}, we define our terms, discuss relevant prior work, and prove some elementary propositions.
+In Section~\ref{sec:commutation} we prove the commutation relation~\ref{eq:commutation-relation}.
+In Section~\ref{sec:reps} we prove another lemma involving representatives for $\A(n)\mmod\A(n-k)$.
+In Section~\ref{sec:li} we prove the key linear independence result.
+Finally, in Section~\ref{sec:lk} we bring the results together to give a sketch of the desired proof method for the general case, and prove the result in some specific small cases.
+
+\section{Definitions and Background}\label{sec:defns}
 
@@
 Recall that the Steenrod squares satisfy the Adem relations:
-\[ \Sq^i\Sq^j = \sum_{m=0}^{m/2}\binom{j-m-1}{i-2m}\Sq^{i+j-k}\Sq^k .\]
+\[ \Sq^i\Sq^j = \sum_{m=0}^{m/2}\binom{j-m-1}{i-2m}\Sq^{i+j-k}\Sq^k.\]
 The binomial coefficient is taken modulo 2, and it is a standard result in combinatorics that it is nonzero if and only if for each bit of the binary representation of $i-2m$ that is 1, the corresponding bit of $j-m-1$ must also be 1.
@@
 \subsection{The Wood \boldmath$Z$ basis}
-In addition to the Adem basis, the $Z$ basis described by Wood~\cite{TODO} will be important to our work.  Let
+In addition to the Adem basis, the $Z$ basis described by Wood~\cite{wood} will be important to our work.  Let
 \[ X_n = \Sq^{2^{n+1}-2^n}\Sq^{2^{n+1}-2^{n-1}}\cdots\Sq^{2^{n+1}-1}, \]
@@
 
-\section{The Commutation Relation}
+\section{The Commutation Relation}\label{sec:commutation}
 
@@
 
-\section{Representatives for \boldmath$\A(n)\mmod\A(n-k)$}
+\section{Representatives for \boldmath$\A(n)\mmod\A(n-k)$}\label{sec:reps}
 
@@
 
-\section{Linear Independence}
+\section{Linear Independence}\label{sec:li}
 
@@
 
-  %TODO: show somewhere that the shortest terms are in fact the shortest, i.e., that all terms of $M_{k, n-k}$ are of length at least $k+1$.
   Note that the terms 
@@
 
-\section{Filtrations of \boldmath$L(k)$}
+\section{Filtrations of \boldmath$L(k)$}\label{sec:lk}
 
@@
 \[ B_{n,k} = \{Sq^I \mid I = (2^{n+1}m_1, 2^n m_2, \ldots, 2^{n-k+2}m_k), m_1 \geq \cdots \geq m_k \} \]
-may allow us to construct a filtration of $\Lkm$ with copies of $\A(n)\mmod\A(n-k)$ as the filtration quotients, through the following theorem.
+may allow us to construct a filtration of $\Lkm$ with copies of $\A(n)\mmod\A(n-k)$ as the filtration quotients, in a similar method to that which Welcher used to prove that $L(k)$ is free over $\A(k-1)$.
 
@@
 
-%TODO: acks
+\section{Acknowledgements}
 
-%TODO: bib
+This paper would not have been possible without the support of Mark Behrens, who proposed the problem and provided guidance along the way, and funding from the MIT Class of 1995 and Paul E.
+Gray Funds for UROP.
+
+\printbibliography
 

### 2015-06-18 paper.tex
maketitle is useful I hear
Add pictures, some other updates

@@
 \author{Benjamin Kraft}
+\date{}
 
@@
 
+\maketitle
+
 \begin{abstract}
-  The Steenrod Algebra $\A$ is the algebra of stable natural endomorphisms of the $\mathbb{Z}/2$-cohomology functor; it is generated by elements $Sq^{2^i}$.
-Let $\A(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.
+  Let $\A$ be the Steenrod Algebra, and let $\A(k)$ be the subalgebra generated by the $Sq^{2^i}$ for $i\leq k$.
 Consider the modules $L(k)$ spanned by sequences of Steenrod operations of length $k$.
@@
 Consider in particular the length filtration of $\A$, with filtration quotients $L(k)$ spanned by sequences of Steenrod operations of length $k$.
+This filtration and its relation to the sphere spectrum has been studied by a number of authors~\cite{nakaoka,welcher,welcher2}.
 
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=1.4in]{pics/A1.pdf}
+    \hspace{1in}
+    \includegraphics[height=1.3in]{pics/A1-A0.pdf}
+  \end{center}
+  \caption{At left, a basis for the subalgebra $\A(1)$.
+The dark lines represent the action of $\Sq^1$ (so for instance $\Sq^1 \cdot \Sq^2\Sq^1 = \Sq^3\Sq^1$), and the light lines represent that of $\Sq^2$.
+At right, representatives for a basis of $\A(1)\mmod\A(0)$.
+Again the dark lines represent the action of $\Sq^1$ and the light lines represent that of $\Sq^2$.\label{fig:A1}}
+\end{figure}
+
 \begin{defn}
@@
 
-%TODO: pictures of small ones
+An example is given in Figure~\ref{fig:A1}.
 In particular, $\A(n)\mmod\A(n-k)$ has as representatives all elements of $\A(n)$ which cannot be formed by multiplying an element of $\A(n)$ by a nontrivial element of $\A(n-k)$.
@@
 
-  %TODO: check this more carefully.
   First, let $y\in L(1)$, $z\in \A(n-k-1)$ with $z$ not in degree $0$, so that $yz\in M_{1,n-k-1}$.
@@
 If we multiply out using the Adem relations, this is clearly true, modulo showing that $zb'$ contains no terms of length $1$.
-This is easy to show; by the Milnor basis, $z$ cannot end in a Steenrod square with degree a multiple of $2^{n-k}$, but $b'$ is a Steenrod square in a degree which is $1$ modulo $2^{n-k}$, so on multiplying the two, the length $1$ term must vanish, as desired.
+This is easy to show; since it is in $\A(n-k-1)$, $z$ cannot end in a Steenrod square with degree a multiple of $2^{n-k}$, but $b'$ is a Steenrod square in a degree which is $1$ modulo $2^{n-k}$, so on multiplying the two, the length $1$ term must vanish, as desired.
 
@@
 Since $\A(n)\mmod\A(n-k)$ is a Poincar\'e algebra, there exists some $y_j$ in $\A(n)\mmod\A(n-k)$ such that $y_j x_j = z$.
-Then for any $b_{j'}$ in a lower degree than $b_j$, the corresponding $x_{j'}$ will be in a higher degree than $x_j$.
+(See for example Margolis~\cite{margolis}.)  Then for any $b_{j'}$ in a lower degree than $b_j$, the corresponding $x_{j'}$ will be in a higher degree than $x_j$.
 Then $y_j x_j = 0$, so multiplying the entire relation by $y_j$ leaves a new relation amongst only the $b_i$ in the highest possible degree.
@@
 
-%TODO: pictures of the filtrations of L(1) and L(2)?
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=2.5in]{pics/L1-1.pdf}
+  \end{center}
+  \caption{The structure of $\L(1)$ as an $\A(1)$-module.
+(As before, the heavy lines show the action of $\Sq^1$, while the lighter lines show the action of $\Sq^2$.)  The colors show each filtration quotient isomorphic to $\A(1)\mmod\A(0)$; the single element $\Sq^1$ is excluded from this, but would be included if we continued downwards to $\Lkm[1]$.
+\label{fig:L1-1}}
+\end{figure}
+
+A picture of the case of $\L(1)$ as an $\A(1)$-module is shown in Figure~\ref{fig:L1-1}.
 
@@
 
-This paper would not have been possible without the support of Mark Behrens, who proposed the problem and provided guidance along the way, and funding from the MIT Class of 1995 and Paul E.
-Gray Funds for UROP.
+This paper would not have been possible without the support of Mark Behrens, who proposed the problem and provided guidance along the way, and sarah-marie belcastro, whose advice during the writing process was invaluable.
+The project was generously funded by the MIT Class of 1995 and Paul E.~Gray Funds for UROP.
 
